---
title: "corpus study"
format: html
editor: visual
---
# WEISST EH

fargene:
https://r-charts.com/colors/

## Setting up workingdirectory and packages

```{r, set up}

# just run this code chunk
# function to ignoring the setting of the relative path below when knitting
run_if_not_knitting <- function(expr) {
  if (!isTRUE(getOption("knitr.in.progress"))) {
    eval(expr)
  }
}

# nifty code using the pacman package
# it checks if the packages specified below are installed, if not, 
#they will be installed, if yes, they will be loaded
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, rstudioapi, tidyverse, bayesrules,
               tidybayes, trimr, brms, bayesplot, bayestestR, modelr, lme4)

# set the current working directory to the one where this file is
run_if_not_knitting(current_working_dir <- dirname(rstudioapi::getActiveDocumentContext()$path))
run_if_not_knitting(setwd(current_working_dir))
```

# Corpus study - loading data and prepare it

```{r, loading data}
#| echo: false 
# loading the data
corpus_raw <- read_delim('tryout.txt', delim = '\t')

corpus <- corpus_raw |>
  mutate(word_form = as.factor(word_form),
         country_corpus = as.factor(country_corpus))
levels(corpus$word_form)

# removing columns not used
#corpus <- corpus



```

Code book:

**full** = complete text of each hit in the corpus including the citation

**word form** = showing the conjugation of the verb and whether the phrase includes the particle *ja* or not

**reference** = the name of the newspaper/original source

**first part of reference** = the abreviation of the original source 

**several hits** = 

**country (corpus)** = 

**speakers nasjonality** = 

**comments** = 

**sentence** = 

**sentence type** = 

**obligatory** = 

**discourse marker** = 

**sentence (position)** = 

**clause (position)** = 

**position** = 

**with conjunction** = 

**conjunction** = 

**in total** =

## descriptive statistics

### word forms across differnt counties

```{r, word forms across differnt counties}
# total number of entries
nrow(corpus)

# count and percentage of different word forms
n_per_wf <- corpus |>
  group_by(word_form) |>
  count()|>
  mutate(percentage = n/nrow(corpus)*100)
print(n_per_wf)

# count and percentage per country
n_per_country <- corpus |>
  group_by(country_corpus) |>
  count()|>
  mutate(percentage = n/nrow(corpus)*100)
print(n_per_country)

# word form counts per country
corpus_wf <- corpus |>
  group_by(word_form, country_corpus)|>
  count()|>
  mutate(percentage = n/nrow(corpus)*100)
print(corpus_wf)

plot_1 <- ggplot(corpus) +
  geom_bar(aes(y = word_form,
               fill = country_corpus),
           stat = "count")+
  facet_wrap(~country_corpus)+
  labs(title = "\n Different word forms found across the different countries",
       y = "")+
  scale_fill_manual(values = c("red", "brown4", "gold"),
                      guide = "none")
  #geom_text(aes(label = n))+
  theme_minimal()
plot_1

ggsave(plot = plot_1, filename = "plot_1_wf_country.png",
       height = 4, width = 6)

# plotting a figure to visualize how the different word forms are distributed across the three countries
plot_2 <- ggplot(corpus_wf,aes(y = reorder(word_form, n,
                                        FUN = function(x) sum(x)),
                            x = n)) +
  geom_bar(stat = "identity",
           aes( fill = country_corpus)) +
  geom_text(aes(label=n,
                colour = country_corpus), size = 3, hjust=-0.2)+
  facet_wrap(~country_corpus)+
  labs(title = "\n Different word forms found across the different countries",
       y = "",
       x = "")+
  scale_fill_manual(values = c("red", "brown4", "#FF8C00"),
                      guide = "none")+
  scale_colour_manual(values = c("#CD6889", "#96CDCD", "peachpuff3"),
                      guide = "none")+
  scale_x_continuous(limits = c(0, 200))+
  #guides(x = guide_axis(angle = 45)) +
  theme_minimal()
# looking at plot
plot_2

# Saving the plot as .png file
ggsave(plot = plot_2, filename = "plot_2_wf_country.png",
       height = 4, width = 6)

```

### position and function

```{r, position and function}

# counting the position
corpus_pos <- corpus |>
  group_by(free_standing) |>
  count() |>
  # adding percentages, rounding to avoid decimals
  mutate(percentage = round(n/nrow(corpus)*100))
print(corpus_pos)
# counting the function
corpus_fun <- corpus |>
  group_by(discourse_marker) |>
  count() |>
  # adding percentages, rounding to avoid decimals
  mutate(percentage = round(n/nrow(corpus)*100))
print(corpus_fun)


# counting how 'weisst eh' is positioned by comparing
# free_standing against discourse_marker
# position:
# free_standing:  0 = embedded in a sentence
#                 1 = sanding completely free as a sentence on its own
# function: 
# discourse_marker: 0 = part of the main clause of the sentence,
#                       not a discourse marker
#                   1 = used as discourse marker
corpus_pos_fun <- corpus |>
  group_by(free_standing, discourse_marker) |>
  count() |>
  # adding percentages, rounding to avoid decimals
  mutate(percentage = round(n/nrow(corpus)*100))
# looking at results
print(corpus_pos)

# adding italics to title
my_title_4 <- expression(paste("Positions and functions of ", italic("weisst eh")))
# plotting the result
plot_4 <- ggplot(corpus_pos_fun,
                 (aes(x = free_standing,
                      y = n)))+
  geom_bar(aes(fill = as.factor(discourse_marker)),
           stat = "identity",
           position = 'dodge')+
  geom_text(aes(label = n,
                colour = as.factor(discourse_marker)),
            position = 'identity',
            size = 3.2,
            vjust=-0.2,
            hjust= c(2.1, -1.2, 5.0, -4.7))+
  scale_fill_manual(values = c("#CD919E","#8B636C"),
                    labels = c("no",
                               "yes"))+
    scale_colour_manual(values = c("#CD919E","#8B636C"))+
  guides(colour = "none")+
  labs(title = my_title_4,
       x = "",
       y = "",
       fill = "used as discourse marker")+
  scale_x_continuous(breaks = c(0,1),
                     labels = c("embedded", "free-standing"))+
  theme_minimal()
plot_4

# Saving the plot as .png file
ggsave(plot = plot_4, filename = "plot_4_position_and_function.png",
       height = 4, width = 6)


```

### funtions across different countries
```{r, functions across country}

# function across countries
corpus_func_count <- corpus |>
  group_by(discourse_marker, country_corpus)|>
  count()
print(corpus_func_count)

# creating a new tibble to plot the results in the optimal way

# get total number of results per country 
print(n_per_country)
# extracting values and saving them in separate objects
AT <- c(n_per_country$n[1])
CH <- c(n_per_country$n[2])
DE <- c(n_per_country$n[3])
# repeating this for the other table
corpus_func_count_n <- c(corpus_func_count$n)
# using the total values per country and order them so that they fit
# together with the corpus_func_count_n object that has 5 lines
total_per_country <- c(n_per_country$n[1],
          n_per_country$n[2],
          n_per_country$n[3],
          n_per_country$n[1],
          n_per_country$n[3])
# combining the two objects above into a new data frame/ tibble
help <- tibble(corpus_func_count_n, total_per_country)
# calculating percentages per country
help <- help |>
  mutate(percentage =
           (corpus_func_count_n/total_per_country))
#|>
  # rounding to avoid decimals
  #mutate(percentage = round(percentage)) |>
 # mutate(percentage = )

help$percentage

# for plottin gwe have to creat a new data frame/tibble
# saving the necessary information in new objects
percentage_per_country <- help$percentage
dis_mar <- corpus_func_count$discourse_marker
country <- corpus_func_count$country_corpus
n <- corpus_func_count$n
# combining those objects into a new data frame/tibble
function_per_country <- tibble(dis_mar,
                               country,
                               n,
                               percentage_per_country)
# looking at the new df
print(function_per_country)
# looks good, then we can plot the results


# adding italics to title
my_title_5 <- expression(paste("Usage of ", italic("weisst eh"), " across different spoken standard varieties"))
# plotting the results
plot_5 <- ggplot(function_per_country,
                 (aes(x = dis_mar,
                      y = n)))+
  geom_bar(aes(fill = as.factor(country)),
           stat = "identity",
           position = 'dodge')+
  geom_text(aes(label = scales::percent(percentage_per_country),
                colour = as.factor(country)),
            position = 'identity',
            size = 3.2,
            vjust=-0.2,
            hjust= c(1,1,1,1,1))+
  #facet_wrap(~discourse_marker)+
  scale_fill_manual(values = c("#CD6889","#96CDCD", "peachpuff3"),
                    labels = c("AT",
                               "CH",
                               "DE"))+
    scale_colour_manual(values = c("#CD6889","#96CDCD", "peachpuff3", "#CD6889", "peachpuff3"))+
  guides(colour = "none")+
  labs(title = my_title_5,
       x = "",
       y = "",
       fill = "standard variety spoken in")+
  scale_x_continuous(breaks = c(0,1),
                     labels = c("main clause", "discourse marker"))+
  theme_minimal()
plot_5

# Saving the plot as .png file
ggsave(plot = plot_5, filename = "plot_5_funcion_per_country.png",
       height = 4, width = 6)

# excluding wikipedia:

# counting how often wikipedia was the source and how that is
# distributed across the function and country and word form
wiki <- corpus |>
  filter(reference == "Wikipedia") |>
  group_by(discourse_marker, country_corpus, word_form )|>
  count()
# only in germany, equally across the two functions
# How do the results look like when we exclude wikipedia
no_wiki <- corpus |>
  filter(reference != "Wikipedia") |>
  group_by(discourse_marker, country_corpus)|>
  count()
# new total of examples when excluding wikipedia (which was n=14)
DE_new <- DE-14
print(DE_new)
# new percentages
# in main clause:
print(round(no_wiki$n[3]/DE_new*100,1))
# as discourse marker
print(round(no_wiki$n[5]/DE_new*100,1))

```


### ja, denn and leicht
```{r, ja, denn and leicht}

# filtering for particle "ja"
corpus_ja <- corpus|>
  filter(ja == 1)

# counting number of examples that had the particle 'ja'
nrow(corpus_ja)

# together with which words forms did 'ja' appear?
corpus_ja |>
  group_by(word_form) |>
  count()
# wasst eh and weisst eh

# In which counties did the corpus search yield results for 'ja'?
corpus_ja |>
  group_by(country_corpus, word_form) |>
  count()
# in all three

# how often was 'ja' used as discourse marker?
corpus_ja |>
  group_by(discourse_marker) |>
  count()
# never

```

### pronouns
```{r, pronouns}

# storing total number of discourse markers in an object
dis_total <- corpus |>
  filter(discourse_marker == 1)|>
  nrow()
# n = 171

# storing total number of "not discourse marker" examples in an object
no_dis_total <- corpus |>
  filter(discourse_marker == 0)|>
  nrow()
# n = 112

# pronouns and their distribution
corpus_pro_dis <- corpus |>
  group_by(with_pronoun, discourse_marker) |>
  count()
# looking at results
print(corpus_pro_dis)
# order of function is: 0 - 1 - 0 - 1 - 0 - 0 

# calculating percentage per function: discourse marker = yes/no
# extracting total number of observation into a new object
corpus_pro_dis_n <- corpus_pro_dis$n
# ordering total numbers of function fitting to the order of
# conditions in corpus_pro_dis_n
total_per_function <- c(no_dis_total, dis_total, no_dis_total, dis_total, no_dis_total, no_dis_total)
# merge into data frame/tibble
help_2 <- tibble(corpus_pro_dis_n, total_per_function)

# calculate percentages:
help_2 <- help_2 |>
  mutate(percentage = corpus_pro_dis_n/total_per_function,
         rounded = round(percentage*100))
# look at result
print(help_2)
```


### conjunctions
```{r, conjunctions}

# how many of total examples have a conjunction right 
# before 'weisst eh' ?
corpus_conj <- corpus |>
  group_by(with_conjunction) |>
  count() |>
  # adding percentage
  mutate(percentage = round((n/nrow(corpus)*100),1))
print(corpus_conj)

# filtering data to only include those examples where 'weisst eh'
# functions as a discourse marker
dismark <- corpus |>
  filter(discourse_marker ==1)

# looking at discourse markers only
dismark_conj <- dismark |>
  group_by(with_conjunction) |>
  count() |>
  mutate(percentage = round((n/nrow(dismark))*100, 1))
print(dismark_conj)

# where is the conjunction placed?
dismark_conj_type <- dismark |>
  group_by(with_conjunction, conjunction)|>
  count()
print(dismark_conj_type)

# comma placement
dismark_conj_type_comma <- dismark |>
  group_by(with_conjunction, conjunction, comma)|>
  count()
print(dismark_conj_type_comma)


```

### discourse markers in detail
```{r, discourse markers}

# filtering out free-standing examples
dismark_emb <- dismark |>
  filter(free_standing == 0)

nrow(dismark_emb)
# 167 examples 

# position with the sentence
dismark_emb_sent <- dismark_emb |>
  group_by(sentence_position)|>
  count() |>
  # adding percentage
  mutate(percentage = round(n/nrow(dismark_emb)*100))
print(dismark_emb_sent)

# position within clause for sentence middle position
dismark_emb_clause <- dismark_emb |>
  # filtering to only use sentence middle examples
  filter(sentence_position == "middle") |>
  group_by(clause_position) |>
  count()
print(dismark_emb_clause)
# conjunction distribution across different positions
dismark_emb_combination_conj <- dismark_emb |>
  group_by(sentence_position, clause_position, with_conjunction)|>
  summarize()
print(dismark_emb_combination_conj)
# pronoun distribution across different positions
dismark_emb_combination_pro <- dismark_emb |>
  group_by(sentence_position, clause_position, with_pronoun)|>
  summarize()
print(dismark_emb_combination_pro)

```



```{r}
explore <- dismark_emb |>
  group_by(conjunction, comments) |>
  summarize()
print(explore)

```






# Exploratory study - Questionnaire

## loading data 

```{r, plots and tables}
# adding italics to title
my_title_5 <- expression(paste("Usage of ", italic("weisst eh"), " across different spoken standard varieties"))
# plotting the results
plot_5 <- ggplot(corpus_func_count,
                 (aes(x = discourse_marker,
                      y = n)))+
  geom_bar(aes(fill = as.factor(country_corpus)),
           stat = "identity",
           position = 'dodge')+
  geom_text(aes(label = n,
                colour = as.factor(discourse_marker)),
            position = 'identity',
            size = 3.2,
            vjust=-0.2,
            hjust= c(4.35,0.6,-3.35,2.45,-2.5))+
  #facet_wrap(~discourse_marker)+
  scale_fill_manual(values = c("#CD6889","#96CDCD", "peachpuff3"),
                    labels = c("AT",
                               "CH",
                               "DE"))+
    scale_colour_manual(values = c("#CD6889","#96CDCD", "peachpuff3", "#CD6889", "peachpuff3"))+
  guides(colour = "none")+
  labs(title = my_title_5,
       x = "",
       y = "",
       fill = "standard variety spoken in")+
  scale_x_continuous(breaks = c(0,1),
                     labels = c("main clause", "discourse marker"))+
  theme_minimal()
plot_5


corpus_func_count|>
  mutate(percentage_per_country = percentage_per_country)
print(corpus_func_count)
```

## analysis

## plotting 

# DELEATE!!!
code junks that I might want t use, but most probably wont need anymore
```{r}
#? kan slettes
corpus |> filter(discourse_marker == 0) |> group_by(with_conjunction) |> count()

corp <- read_delim('correct_spelling.txt', delim = '\t')
# word form counts per country
corp_wf <- corp |>
  group_by(word_form, country_corpus)|>
  count()
corp
print(corp_wf)

plot_1 <- ggplot(corpus_wf) +
  geom_(aes(y = word_form,
               fill = country_corpus,))+
  labs(title = "\n Different word forms found across the different countries",
       y = "")+
  theme_minimal()
plot_1

,
       y = c( "waßt eh", "waßt ja eh", "weißt eh", "weißt ja eh", "weisst eh", "weisst ja eh", "wissen's eh", "wisst's eh", "woaßt eh", "woaßt ja eh")
              
# adjusting levels of word_form column
corpus_reo <- as.character(corpus$word_form,
                       levels = c("weisst eh", "wasst eh", "woasst eh", "wisst's eh", "wissen's eh"))

# adjusting columns for analysis
corpus_adj <- corpus_raw |>
  mutate(
    word_form =
      ifelse(
        original_word_form == "wa<df>t eh"~"wasst eh",
        "wa<df>t ja eh"~"wasst eh",
        "weisst eh"~"weisst eh",
        "weisst ja eh"~"weisst eh",
        "wei<df>t eh"~"weisst eh",
        "wei<df>t ja eh"~"weisst eh",
        "wissen's eh" ~"wissen's eh",
        "wisst's eh"~"wisst's eh",
        "woasst eh"~"woasst eh",
        "woa<df>t eh"~"woasst eh"))
    complete_word_form_adjusted =
      ifelse(
        original_word_form == "wa<df>t eh"~"wabt eh",
        "wa<df>t ja eh"~"wabt ja eh",
        "weisst eh"~"weisst eh",
        "weisst ja eh"~"weisst ja eh",
        "wei<df>t eh"~"weibt eh",
        "wei<df>t ja eh"~"weibt ja eh",
        "wissen's eh" ~"wissen's eh",
        "wisst's eh"~"wisst's eh",
        "woasst eh"~"woasst eh",
        "woa<df>t eh"~"woabt eh"),
    ja =
      ifelse(
        original_word_form == "wa<df>t eh"~0,
        "wa<df>t ja eh"~1,
        "weisst eh"~0,
        "weisst ja eh"~1,
        "wei<df>t eh"~0,
        "wei<df>t ja eh"~1,
        "wissen's eh" ~0,
        "wisst's eh"~0,
        "woasst eh"~0,
        "woa<df>t eh"~0)
      )
```
```{r, copy of above in case i need it}
# counting how 'weisst eh' is positioned by comparing
# free_standing against discourse_marker

# free_standing:  0 = embedded in a sentence
#                 1 = sanding completely free as a sentence on its own
# discourse_marker: 0 = part of the main clause of the sentence,
#                       not a discourse marker
#                   1 = used as discourse marker
corpus_pos <- corpus |>
  group_by(free_standing, discourse_marker) |>
  count() |>
  # adding percentages, rounding to avoid decimals
  mutate(percentage = round(n/nrow(corpus)*100))
# looking at results
print(corpus_pos)

# create a new column that combines free_standing and discourse_marker
corpus_pos <- corpus_pos |>
  mutate(comb_free_disk = case_when(free_standing == 0 &
                                  discourse_marker == 0 ~ "00",
                                free_standing == 0 &
                                  discourse_marker == 1 ~ "01",
                                free_standing == 1 &
                                  discourse_marker == 0 ~ "10",
                                free_standing == 1 &
                                  discourse_marker == 1 ~ "11",
                                )) |>
  mutate(position = case_when(free_standing == 0 ~ "embedded",
                                   free_standing == 1 ~"free standing"))

  # removing columns we don't need
#corpus_pos <- corpus_pos |> select(-free_standing, -discourse_marker)
# looking at result
corpus_pos

# plotting the result
plot_3 <- ggplot(corpus_pos,
                 aes(x = "",
                     y = n,
                     fill = comb_free_disk))+
  geom_col()+
  geom_text(aes(label = n),
            position = position_stack(vjust = 0.6))+
  coord_polar(theta = "y")+
  scale_fill_manual(values =c ("#8B7D7B","#EED5D2","#AEEEEE",  "#96CDCD"))+
  theme_void()
plot_3

corpus_pos <- corpu
plot_4 <- ggplot(corpus_pos,
                 (aes(x = free_standing,
                      y = n)))+
  geom_bar(aes(fill = as.factor(discourse_marker)),
           stat = "identity",
           position = 'dodge')+
  geom_text(aes(label = n),
                #colour = as.factor(discourse_marker)),
            position = 'identity', size = 3, vjust=-0.2,
            guide = "none")+
  scale_fill_manual(values = c("#CD919E","#8B636C"),
                    labels = c("no",
                               "yes"))+
    scale_colour_manual(values = c("#CD919E","#8B636C"))+
  labs(title = "Positions and functions of 'weisst eh' ",
       x = "",
       y = "",
       fill = "used as discourse marker")+
  scale_x_continuous(breaks = c(0,1),
                     labels = c("embedded", "free standing"))+
  theme_minimal()
plot_4

# fix the numbers




"#AEEEEE",  "#96CDCD"
```
### destriptive statistics III copy
```{r, descriptive statistics III copy}

# position and function across countries
corpus_pos_func_count <- corpus |>
  group_by(free_standing, discourse_marker, country_corpus)|>
  count()

print(corpus_pos_func_count)


plot_5 <- ggplot(corpus_pos_func_count,
                 (aes(x = free_standing,
                      y = n)))+
  geom_bar(aes(fill = as.factor(country_corpus)),
           stat = "identity",
           position = 'dodge')+
  geom_text(aes(label = n,
                colour = as.factor(discourse_marker)),
            position = 'identity',
            size = 3.2,
            vjust=-0.2,
            hjust= c(1))+
  facet_wrap(~discourse_marker)+
  scale_fill_manual(values = c("#CD6889","#96CDCD", "peachpuff3"),
                    labels = c("AT",
                               "CH",
                               "DE"))+
    scale_colour_manual(values = c("#CD6889","#96CDCD", "peachpuff3"))+
  guides(colour = "none")+
  labs(title = my_title,
       x = "",
       y = "",
       fill = "used as discourse marker")+
  scale_x_continuous(breaks = c(0,1),
                     labels = c("embedded", "free-standing"))+
  theme_minimal()
plot_5
```
